A finite-state morphological transducer for Kyrgyz

### 2011-10-14 dev/paper/paper.tex
and moar paper

\usepackage{booktabs}
\usepackage{natbib}

@@

@@

@@

>>>>

\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{natbib}

@@

%TODO: for a future version
%\subsection{Challenges:  Copula, how to represent different tenses, what is a “word”?}

%FIXME: do we want this?
\subsection{Irregular negatives of finite verb forms}

@@

%FIXME: do we want this?
\subsection{Irregular [noun $+$ possessive $+$ case] forms}

@@

\section{Morphophonology}

%FIXME: do we want this?
\subsection{Vowel harmony}

In Kyrgyz, there are two basic archiphonemes: a low vowel, \{A\}, and a high vowel, \{I\}.  The high vowel takes the backness and rounding of the preceding vowel, resulting in the values shown in table \ref{vowelharmonyI}.

\begin{table}[htbp]
	\centering
	\caption{vowel harmony for archiphoneme \{I\}}\label{vowelharmonyI}
	\subfloat{
		\begin{tabular}{ll}
			\toprule
			after & result \\
			\midrule
			и & и \\
			ү & ү \\
			е & и \\
			ө & ү \\
			\bottomrule
		\end{tabular}
	}
	\qquad
	\subfloat{
		\begin{tabular}{ll}
			\toprule
			after & result \\
			\midrule
			ы & ы \\
			у & у \\
			а & ы \\
			о & у \\
			\bottomrule
		\end{tabular}
	}
\end{table}

The back vowel \{A\} also takes the backness and rounding of the preceding vowel, with the exception of when it occurs after /у/---n this case, it has an unrounded variant, as depicted in table \ref{vowelharmonyA}.

\begin{table}[htbp]
	\centering
	\caption{vowel harmony for archiphoneme \{A\}}\label{vowelharmonyA}
	\subfloat{
		\begin{tabular}{ll}
			\toprule
			after & result \\
			\midrule
			и & е \\
			ү & ө \\
			е & е \\
			ө & ө \\
			\bottomrule
		\end{tabular}
	}
	\qquad
	\subfloat{
		\begin{tabular}{ll}
			\toprule
			after & result \\
			\midrule
			ы & а \\
			у & а \\
			а & а \\
			о & о \\
			\bottomrule
		\end{tabular}
	}
\end{table}

There were other vowel archiphonemes that had to be implemented in this project.  For example, \{U\} occurs in the past habitual suffix -/чU/ and the general gerund/infinitive -/Uː/, but nowhere else in the language.  Also, \{E\} was created for use in the habitual/future suffix; it has surface forms identical to those of \{A\}, except after vowels where it surfaces as [й].  Despite the fact that \{E\} is very similar to \{A\}, a separate but very similar twolc rule had to be created.  Repeated content could have been reduced by using cascading rules instead of two-level rules, but this would have caused other complications, such as finding the correct rule ordering.  An alternative possibility would’ve been to have a single intermediate level so that a rule like ``\{E\}→\{A\} after vowels'' with a default surface form of [й] could've been implemented.

%FIXME: do we want this?
\subsection{Consonant-consonant phonology: desonorisation and assimilation}
In Kyrgyz, there are two basic processes affecting the realisation of consonants when they are adjacent to other consonants: voicing assimilation and desonorisation.

Voicing assimilation in Kyrgyz involves the agreement of voicing of two consonants across a syllable boundary.  An example involves the locative and dative suffixes, as shown in table \ref{exvoicing}.
